Future Evolution of CSMA Protocols for the IEEE 802.11 Standard

### 2013-03-01 paper.tex
playing with colors

When it comes to backward compatibility, neither of the aforementioned solutions are able to fairly share the medium with legacy devices.
In fact, these proposals are, generally speaking, less aggressive than the currently implemented CSMA/CA protocol. Consequently, in a hypothetical mixed network in which the new and old protocols coexist, the new stations would receive a smaller share of the available bandwidth.

@@

Quality-of-service and traffic differentiation in CSMA/ECA has is discussed in \cite{barcelo2009tpc, he2009srb}.
The performance of CSMA/ECA in the presence of channel errors is considered in \cite{he2009srb, barcelo2010fcc, fang2011dlm, barcelo2012mdc}. 
Other aspects such as coexistence with legacy stations and performance in the presence of new entrants have also been covered in \cite{he2009srb, barcelo2010fcc, fang2011dlm}.
Results for the impact of slot drift (non-ideal clocks) in protocols similar to CSMA/ECA can be found in \cite{gong2012asd}.
In the present workshop paper, we focus on a simple greenfield scenario with a single traffic class, ideal channel, ideal clocks, and saturation conditions.
We are currently evaluating more complex scenarios that we outline in Section \ref{sec:}

@@

In the current paper, we present a completely distributed solution, called {\it backoff stage hysteresis} or simply hysteresis, which is a simple modification to CSMA/ECA to accommodate a large number of contenders in an adaptive and distributed fashion. Hysteresis by itself, comes at the expense of reduction in long-term fairness. Using hysteresis together with what we will refer to as {\it fare-share} allows for reaching collision-free operation with self-adaptation to a variable, and potentially large, number of contenders while preserving fairness. The concept of fair-share was first introduced in \cite{fang2011dlm}. We will furthermore detail pseudo-code algorithms for three different types of CSMA/ECA (basic, with hysteresis, and with both hysteresis and fair-share) and highlight how they can be easily implemented using simple modifications to the currently-used CSMA/CA algorithm. We will then present some performance evaluation, quantifying the performance gain that these three types of CSMA/ECA can achieve, both in terms of throughput and fairness.

@@

\begin{algorithm}[ht!!!]
\While{the device is on}
{
  %\tcc{Initialize retransmission attempts.}
  $r \leftarrow 0$ ; $s \leftarrow 0$ \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission of 1 packet \;
      \If{collision}{
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min (s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min} -1]$\;
      }
    }
    $r \leftarrow 0$ \;
    $s \leftarrow 0$ \;
	 \eIf{success}{
      %\tcc{Random backoff.}
      $b \leftarrow \mathcal{U}[0,2^{s}\rm{CW}_{min}-1]$\;
    }
    {
      Discard packet\;
      $b \leftarrow \mathcal{U}[0,2^s \rm{CW}_{min}-1]$\;
    }
  }
  Wait until there is a packet to transmit \;
}
\caption{CSMA/CA}
\label{alg:CSMA_CA}
\end{algorithm}

@@

\section{Future Work}

>>>>

When it comes to backward compatibility, neither of the aforementioned solutions are able to fairly share the medium with legacy devices.
In fact, these proposals are, generally speaking, less aggressive than the currently implemented CSMA/CA protocol. Consequently, in a hypothetical mixed network in which the new and old protocols coexisted, the new stations would receive a smaller share of the available bandwidth.

@@

Quality-of-service and traffic differentiation in CSMA/ECA has is discussed in \cite{barcelo2009tpc, he2009srb}.
The performance of CSMA/ECA in the presence of channel errors is considered in \cite{he2009srb, barcelo2010fcc, fang2011dlm, barcelo2012mdc}. 
Other aspects such as coexistence with legacy stations and performance in the presence of new entrants have also been covered in \cite{he2009srb, barcelo2010fcc, fang2011dlm}.
Results for the impact of slot drift (non-ideal clocks) in protocols similar to CSMA/ECA can be found in \cite{gong2012asd}.
In the present workshop paper, we focus on a simple greenfield scenario with a single traffic class, ideal channel, ideal clocks, and saturation conditions.
We are currently evaluating more complex scenarios that we outline in Section \ref{sec:future_work}

@@

In the current paper, we present a completely distributed solution, called {\it backoff stage hysteresis} or simply hysteresis, which is a simple modification to CSMA/ECA to accommodate a large number of contenders in an adaptive and distributed fashion. Hysteresis by itself, comes at the expense of reduction in long-term fairness. Using hysteresis together with what we will refer to as {\it fair-share} allows for reaching collision-free operation with self-adaptation to a variable, and potentially large, number of contenders while preserving fairness. The concept of fair-share was first introduced in \cite{fang2011dlm}. We will furthermore detail pseudo-code algorithms for three different types of CSMA/ECA (basic, with hysteresis, and with both hysteresis and fair-share) and highlight how they can be easily implemented using simple modifications to the currently-used CSMA/CA algorithm. We will then present some performance evaluation, quantifying the performance gain that these three types of CSMA/ECA can achieve, both in terms of throughput and fairness.

@@

\begin{algorithm}[ht!!!]
\While{the device is on}
{
  %\tcc{Initialize retransmission attempts.}
  \colorbox{red}{$r \leftarrow 0$ ; $s \leftarrow 0$} \;
  %\tcc{Initialize backoff counter.}
  $b \leftarrow \mathcal{U}[0,2^s\rm{CW}_{min}-1]$\;
  \While{there is a packet to transmit}{
    %\tcc{Initialize $a$.}
    \Repeat{($r = R$) or (success)}{
      %\tcc{First, backoff.}
      \While{$b>0$}{
        wait 1 slot \;
        $b \leftarrow b-1$ \;
      }
      Attempt transmission of 1 packet \;
      \If{collision}{
        %\tcc{Random backoff.}
        $r \leftarrow r+1$ \;
        $s \leftarrow \min (s+1,S)$ \;
        $b \leftarrow \mathcal{U}[0, 2^s \rm{CW}_{min} -1]$\;
      }
    }
    $r \leftarrow 0$ \;
    $s \leftarrow 0$ \;
	 \eIf{success}{
      %\tcc{Random backoff.}
      $b \leftarrow \mathcal{U}[0,2^{s}\rm{CW}_{min}-1]$\;
    }
    {
      Discard packet\;
      $b \leftarrow \mathcal{U}[0,2^s \rm{CW}_{min}-1]$\;
    }
  }
  Wait until there is a packet to transmit \;
}
\caption{CSMA/CA}
\label{alg:CSMA_CA}
\end{algorithm}

@@

\section{Future Work}
\label{sec:future_work}
